Exploratory Analysis Ethan Schilling

### 2025-12-10 Exploratory Analysis Ethan SchillingREAL.Rmd
Added table of results for linear models predicting on FeCl_dose

getwd()
```

```{r}
library(lubridate)
library(dplyr)
library(ggplot2)
library(tidyr)
library(Metrics)

@@

# 1 Interaction Term
model1 <- lm(FeCl_dose ~ mean_Temp * mean_Turbidity + mean_pH + Temp_lag1 + Turb_lag1, 
             data = train)
summary(model1)

@@

rmse(test$FeCl_dose, test$pred_model1)
mae(test$FeCl_dose, test$pred_model1)

# 2 Interaction Terms
model2 <- lm(FeCl_dose ~ mean_Temp * mean_pH +mean_Conductivity * mean_Turbidity + Temp_lag1 + Temp_lag2 + pH_lag1 + pH_lag2, data = train)

@@

rmse(test$FeCl_dose, test$pred_model2)
mae(test$FeCl_dose, test$pred_model2)

# Step Up

response_var <- "FeCl_dose" 
predictors <- c("mean_Temp", "mean_pH", "mean_Conductivity",
                "mean_Turbidity", "Temp_lag1", "Temp_lag2",
                "pH_lag1", "pH_lag2", "Cond_lag1", "Cond_lag2",
                "Turb_lag1", "Turb_lag2")

>>>>

```{r}
library(lubridate)
library(dplyr)
library(ggplot2)
library(tidyr)
library(Metrics)
library(kableExtra)
library(gt)

@@

# All predictors

model0 <- lm(FeCl_dose ~ mean_Temp + mean_pH + mean_Conductivity + mean_Turbidity + Temp_lag1 + Temp_lag2 + pH_lag1 + pH_lag2 + Cond_lag1 + Cond_lag2 + Turb_lag1 + Turb_lag2, data = train)

test$pred_model0 <- predict(model0, newdata = test)

# 1 Interaction Term

model1 = lm(FeCl_dose ~ mean_Temp * mean_pH + mean_Turbidity + Temp_lag1  + Temp_lag2 + Turb_lag1 +     Turb_lag2,data = train)
summary(model1)

@@

# 2 Interaction Terms

model2 = lm(FeCl_dose ~ mean_Temp * mean_pH + mean_Conductivity * mean_Turbidity +
                Temp_lag1 + Temp_lag2 + pH_lag1 + pH_lag2 + Cond_lag1 + Cond_lag2, data = train)

@@

model3 = lm(FeCl_dose ~ mean_Temp * mean_Conductivity + Temp_lag1 + Temp_lag2 + Cond_lag1 + Cond_lag2,                 data = train)

test$pred_model3 <- predict(model3, newdata = test)

# 2 Interactions: Cond and PH and Temp and Cond

model4 = lm(FeCl_dose ~ mean_Conductivity * mean_pH + mean_Temp * mean_Conductivity +
                Temp_lag1 + Temp_lag2 + pH_lag1 + pH_lag2 + Cond_lag1 + Cond_lag2,
              data = train)

test$pred_model4 <- predict(model4, newdata = test)

model5 = lm(FeCl_dose ~ mean_Temp + mean_pH + Temp_lag1 + Temp_lag2 + pH_lag1 + pH_lag2,
              data = train)

test$pred_model5 <- predict(model5, newdata = test)

models <- list(
  model0 = lm(FeCl_dose ~ mean_Temp + mean_pH + mean_Conductivity + mean_Turbidity +
                Temp_lag1 + Temp_lag2 + pH_lag1 + pH_lag2 + Cond_lag1 + Cond_lag2 +
                Turb_lag1 + Turb_lag2, data = train),

  model1 = lm(FeCl_dose ~ mean_Temp * mean_pH + mean_Turbidity + Temp_lag1  + Temp_lag2 + Turb_lag1 +
                Turb_lag2,data = train),

  model2 = lm(FeCl_dose ~ mean_Temp * mean_pH + mean_Conductivity * mean_Turbidity +
                Temp_lag1 + Temp_lag2 + pH_lag1 + pH_lag2 + Cond_lag1 + Cond_lag2, data = train),

  model3 = lm(FeCl_dose ~ mean_Temp * mean_Conductivity + Temp_lag1 + Temp_lag2 + Cond_lag1 + Cond_lag2,                 data = train),

  model4 = lm(FeCl_dose ~ mean_Conductivity * mean_pH + mean_Temp * mean_Conductivity +
                Temp_lag1 + Temp_lag2 + pH_lag1 + pH_lag2 + Cond_lag1 + Cond_lag2,
              data = train),

  model5 = lm(FeCl_dose ~ mean_Temp + mean_pH + Temp_lag1 + Temp_lag2 + pH_lag1 + pH_lag2,
              data = train)
)

descriptions <- c(
  "Main effects only",
  "Temp × pH interaction",
  "Temp × pH + Cond × Turb interactions",
  "Temp × Cond interaction",
  "Cond × pH + Temp × Cond interactions",
  "Only Temp and pH"
)

performance <- data.frame(Model = character(),
                          RMSE = numeric(),
                          MAE = numeric(),
                          Description = character(),
                          stringsAsFactors = FALSE)

for(i in 1:length(models)) {
  pred <- predict(models[[i]], newdata = test)
  performance <- rbind(performance,
                       data.frame(Model = paste0("Model ", i-1),
                                  RMSE = round(rmse(test$FeCl_dose, pred), 3),
                                  MAE  = round(mae(test$FeCl_dose, pred), 3),
                                  Description = descriptions[i],
                                  stringsAsFactors = FALSE))

}
  performance <- performance %>% arrange(RMSE)

best_rmse <- min(performance$RMSE)

kable_table <- performance %>%
  kable("html", caption = "Model Performance Summary") %>%
  kable_styling(full_width = F, position = "center") %>%
  row_spec(which(performance$RMSE == best_rmse), background = "#cfe2f3")

kable_table
